---
title: "Leaflet intro from 'Leaflet for R' web page - 2"
subtitle: "part 2 : Markers, popup, labels"
author: "Bruno Fischer Colonimos"
date: "16 septembre 2017"
output: 
    html_document:
        toc: yes
        toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install packages if needed
```{r}
# install.packages("leaflet") #**NO ==> IS NOT OK
# When compiling to HTML, standard tiles do not show
# instead:
# devtools::install_github('rstudio/leaflet')
# 
# install.packages("sp")
# install.packages("maps")
# install.packages("rgeos")
# install.packages("htmltools")
```



Loading the package(s)
```{r}
library(leaflet)
```


Markers
=======
Use markers to call out points on the map. Marker locations are expressed in latitude/longitude coordinates, and can either appear as icons or as circles.

Data sources
------------
Point data for markers can come from a variety of sources:

    SpatialPoints or SpatialPointsDataFrame objects (from the sp package)
    POINT, sfc_POINT, and sf objects (from the sf package); only X and Y dimensions will be considered
    Two-column numeric matrices (first column is longitude, second is latitude)
    Data frame with latitude and longitude columns. You can explicitly tell the marker function which columns contain the coordinate data (e.g. addMarkers(lng = ~Longitude, lat = ~Latitude)), or let the function look for columns named lat/latitude and lon/lng/long/longitude (case insensitive).
    Simply provide numeric vectors as lng and lat arguments

Note that MULTIPOINT objects from sf are not supported at this time.

Icon Markers
------------
Icon markers are added using the addMarkers or the addAwesomeMarkers functions. Their default appearance is a dropped pin. As with most layer functions, the popup argument can be used to add a message to be displayed on click, and the label option can be used to display a text label either on hover or statically.

```{r}
data(quakes)
# head(quakes)

# Show first 20 rows from the `quakes` dataset
leaflet(data = quakes[1:20,]) %>% 
        addTiles() %>%
        addMarkers(~long, ~lat, popup = ~as.character(mag), 
                   label = ~as.character(mag))
```

Customizing Marker Icons
------------------------
You can provide custom markers in one of several ways, depending on the scenario. For each of these ways, the icon can be provided as either a URL or as a file path.

For the simple case of applying a single icon to a set of markers, use makeIcon().

```{r}
greenLeafIcon <- makeIcon(
  iconUrl = "http://leafletjs.com/examples/custom-icons/leaf-green.png",
  iconWidth = 38, iconHeight = 95,
  iconAnchorX = 22, iconAnchorY = 94,
  shadowUrl = "http://leafletjs.com/examples/custom-icons/leaf-shadow.png",
  shadowWidth = 50, shadowHeight = 64,
  shadowAnchorX = 4, shadowAnchorY = 62
)

leaflet(data = quakes[1:4,]) %>% addTiles() %>%
  addMarkers(~long, ~lat, icon = greenLeafIcon)
```

If you have several icons to apply that vary only by a couple of parameters (i.e. they share the same size and anchor points but have different URLs), use the icons() function. icons() performs similarly to data.frame(), in that any arguments that are shorter than the number of markers will be recycled to fit.

```{r}
quakes1 <- quakes[1:10,]

leafIcons <- icons(
  iconUrl = ifelse(quakes1$mag < 4.6,
    "http://leafletjs.com/examples/custom-icons/leaf-green.png",
    "http://leafletjs.com/examples/custom-icons/leaf-red.png"
  ),
  iconWidth = 38, iconHeight = 95,
  iconAnchorX = 22, iconAnchorY = 94,
  shadowUrl = "http://leafletjs.com/examples/custom-icons/leaf-shadow.png",
  shadowWidth = 50, shadowHeight = 64,
  shadowAnchorX = 4, shadowAnchorY = 62
)

leaflet(data = quakes1) %>% addTiles() %>%
  addMarkers(~long, ~lat, icon = leafIcons)
```

Finally, if you have a set of icons that vary in multiple parameters, it may be more convenient to use the iconList() function. It lets you create a list of (named or unnamed) makeIcon() icons, and select from that list by position or name.


```{r}
# Make a list of icons. We'll index into it based on name.
oceanIcons <- iconList(
  ship = makeIcon("ferry-18.png", "ferry-18@2x.png", 18, 18),
  pirate = makeIcon("danger-24.png", "danger-24@2x.png", 24, 24)
)

# Some fake data
df <- sp::SpatialPointsDataFrame(
  cbind(
    (runif(20) - .5) * 10 - 90.620130,  # lng
    (runif(20) - .5) * 3.8 + 25.638077  # lat
  ),
  data.frame(type = factor(
    ifelse(runif(20) > 0.75, "pirate", "ship"),
    c("ship", "pirate")
  ))
)

leaflet(df) %>% addTiles() %>%
  # Select from oceanIcons based on df$type
  addMarkers(icon = ~oceanIcons[type])
```

Awesome Icons
-----------------
Leaflet supports even more customizable markers using the awesome markers leaflet plugin.

The `addAwesomeMarkers()` function is similar to `addMarkers()` function but additionally allows you to specify custom colors for the markers as well as icons from the Font Awesome, Bootstrap Glyphicons, and Ion icons icon libraries.

Similar to the makeIcon, icons, and iconList functions described above, you have makeAwesomeIcon, awesomeIcons and awesomeIconList functions, which enable you to add awesome icons.


```{r}
# first 20 quakes
df.20 <- quakes[1:20,]

getColor <- function(quakes) {
        sapply(quakes$mag, function(mag) {
                if(mag <= 4) {
                        "green"
                } else if(mag <= 5) {
                        "orange"
                } else {
                        "red"
                } })
}

icons <- awesomeIcons(
        icon = c('info', 'line-chart', 'map'),
        iconColor = 'black',
        library = 'fa',
        markerColor = getColor(df.20),
        # iconRotate = 45
        spin = c(TRUE, FALSE, FALSE),
        squareMarker = TRUE
)

leaflet(df.20) %>% addTiles() %>%
        addAwesomeMarkers(~long, ~lat, icon=icons, label=~as.character(mag))
```


The library argument has to be one of ‘ion’, ‘fa’, or ‘glyphicon’. The icon argument needs to be the name of any valid icon supported by the the respective library (w/o the prefix of the library name).

Marker Clusters
----------------
When there are a large number of markers on a map, you can cluster them using the Leaflet.markercluster plug-in. To enable this plug-in, you can provide a list of options to the argument clusterOptions, e.g.

```{r}
leaflet(quakes) %>% addTiles() %>% addMarkers(
  clusterOptions = markerClusterOptions()
)
```

Using the freezeAtZoom argument of the markerClusterOptions() function you can set the clustering to freeze as a specific zoom level. For example markerClusterOptions(freezeAtZoom = 5) will freeze the cluster at zoom level 5 regardless of the user’s actual zoom level.

```{r}
leaflet(quakes) %>% addTiles() %>% addMarkers(
  clusterOptions = markerClusterOptions(freezeAtZoom = 5)
)
```




Circle Markers
--------------
Circle markers are much like regular circles (see Lines and Shapes), except that their radius in onscreen pixels stays constant regardless of zoom level.

You can use their default appearance:
```{r}
leaflet(df) %>% 
        addTiles() %>% 
        addCircleMarkers()
```

Or customize their color, radius, stroke, opacity, etc.

```{r}
# Create a palette that maps factor levels to colors
pal <- colorFactor(c("navy", "red"), domain = c("ship", "pirate"))

leaflet(df) %>% addTiles() %>%
  addCircleMarkers(
    radius = ~ifelse(type == "ship", 6, 10),
    color = ~pal(type),
    stroke = FALSE, fillOpacity = 0.5
  )

```


Popups
========
Popups are small boxes containing arbitrary HTML, that point to a specific point on the map.

Use the addPopups() function to add standalone popup to the map.

```{r}
content <- paste(sep = "<br/>",
  "<b><a href='http://www.samurainoodle.com'>Samurai Noodle</a></b>",
  "606 5th Ave. S",
  "Seattle, WA 98138"
)

leaflet() %>% addTiles() %>%
  addPopups(-122.327298, 47.597131, content,
    options = popupOptions(closeButton = FALSE)
  )
```

A common use for popups is to have them appear when markers or shapes are clicked. Marker and shape functions in the Leaflet package take a popup argument, where you can pass in HTML to easily attach a simple popup.

```{r}
library(htmltools)

df <- read.csv(textConnection(
"Name,Lat,Long
Samurai Noodle,47.597131,-122.327298
Kukai Ramen,47.6154,-122.327157
Tsukushinbo,47.59987,-122.326726"
))

leaflet(df) %>% addTiles() %>%
  addMarkers(~Long, ~Lat, popup = ~htmlEscape(Name))
```

In the preceding example, htmltools::htmlEscape was used to santize any characters in the name that might be interpreted as HTML. While it wasn’t necessary for this example (as the restaurant names contained no HTML markup), doing so is important in any situation where the data may come from a file or database, or from the user.

In addition to markers you can also add popups on shapes like lines, circles and other polygons.


Labels
======
A label is a textual or HTML content that can attached to markers and shapes to be always displayed or displayed on mouse over. Unlike popups you don’t need to click a marker/polygon for the label to be shown.

```{r}
library(htmltools)

df <- read.csv(textConnection(
"Name,Lat,Long
Samurai Noodle,47.597131,-122.327298
Kukai Ramen,47.6154,-122.327157
Tsukushinbo,47.59987,-122.326726"))

leaflet(df) %>% 
        addTiles() %>%
        addMarkers(~Long, ~Lat, label = ~htmlEscape(Name))
```

Customizing Marker Labels
-------------------------
You can customize marker labels using the labelOptions argument of the addMarkers function. The labelOptions argument can be populated using the labelOptions() function. If noHide is false (the default) then the label is displayed only when you hover the mouse over the marker; if noHide is set to true then the label is always displayed.

```{r}
# Change Text Size and text Only and also a custom CSS
leaflet() %>% 
        addTiles() %>% 
        setView(-118.456554, 34.09, 13) %>%
        addMarkers(
                lng = -118.456554, lat = 34.105,
                label = "Default Label",
                labelOptions = labelOptions(noHide = T)) %>%
        addMarkers(
                lng = -118.456554, lat = 34.095,
                label = "Label w/o surrounding box",
                labelOptions = labelOptions(noHide = T, 
                                            textOnly = TRUE)) %>%
        addMarkers(
                lng = -118.456554, lat = 34.085,
                label = "label w/ textsize 15px",
                labelOptions = labelOptions(noHide = T, 
                                            textsize = "15px")) %>%
        addMarkers(
                lng = -118.456554, lat = 34.075,
                label = "Label w/ custom CSS style",
                labelOptions = labelOptions(noHide = T, 
                                            direction = "bottom",
                                            style = list(
                                                    "color" = "red",
                                                    "font-family" = "serif",
                                                    "font-style" = "italic",
                                                    "box-shadow" = "3px 3px rgba(0,0,0,0.25)",
                                                    "font-size" = "12px",
                                                    "border-color" = "rgba(0,0,0,0.5)"
                                            )))
```


Labels without markers
-----------------------
You can create labels without the accompanying markers using the addLabelOnlyMarkers function.



