---
title: "Leaflet intro from 'Leaflet for R' web page"
author: "Bruno Fischer Colonimos"
date: "16 septembre 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install packages if needed
```{r}
# install.packages("leaflet")
# install.packages("sp")
# install.packages("maps")
install.packages("rgeos")
```



Loading the package(s)
```{r}
library(leaflet)
```


Introduction
=============
Leaflet is one of the most popular open-source JavaScript libraries for interactive maps. It’s used by websites ranging from The New York Times and The Washington Post to GitHub and Flickr, as well as GIS specialists like OpenStreetMap, Mapbox, and CartoDB.

This R package makes it easy to integrate and control Leaflet maps in R.

Features
--------
    Interactive panning/zooming
    Compose maps using arbitrary combinations of:
        Map tiles
        Markers
        Polygons
        Lines
        Popups
        GeoJSON
    Create maps right from the R console or RStudio
    Embed maps in knitr/R Markdown documents and Shiny apps
    Easily render spatial objects from the sp or sf packages, or data frames with latitude/longitude columns
    Use map bounds and mouse events to drive Shiny logic
    Display maps in non spherical mercator projections
    Augment map features using chosen plugins from leaflet plugins repository


Installation
--------------
To install this R package, run this command at your R prompt:

```{r}
# install.packages("leaflet")
# to install the development version from Github, run
# devtools::install_github("rstudio/leaflet")
```

Once installed, you can use this package at the R console, within R Markdown documents, and within Shiny applications.

Basic Usage
-------------
You create a Leaflet map with these basic steps:

* Create a map widget by calling leaflet().
* Add layers (i.e., features) to the map by using layer functions (e.g. addTiles, addMarkers, addPolygons) to modify the map widget.
* Repeat step 2 as desired.
* Print the map widget to display it.


Here’s a basic example:

```{r}
library(leaflet)

m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=174.768, lat=-36.852, popup="The birthplace of R")
m  # Print the map

```


Next Steps
----------
We highly recommend that you proceed to The Map Widget page before exploring the rest of this site, as it describes common idioms we’ll use throughout the examples on the other pages.

Although we have tried to provide an R-like interface to Leaflet, you may want to check out the API documentation of Leaflet occasionally when the meanings of certain parameters are not clear to you.





The Map Widget
==============

The Data Object
---------------

Both leaflet() and the map layer functions have an optional data parameter that is designed to receive spatial data in one of several forms:

* From base R:
    * lng/lat matrix
    * data frame with lng/lat columns
* From the sp package:
    * SpatialPoints[DataFrame]
    * Line/Lines
    * SpatialLines[DataFrame]
    * Polygon/Polygons
    * SpatialPolygons[DataFrame]
* From the maps package:
    * the data frame from returned from map()

The data argument is used to derive spatial data for functions that need it; for example, if data is a SpatialPolygonsDataFrame object, then calling addPolygons on that map widget will know to add the polygons from that SpatialPolygonsDataFrame.

It is straightforward to derive these variables from sp objects since they always represent spatial data in the same way. On the other hand, for a normal matrix or data frame, any numeric column could potentially contain spatial data. So we resort to guessing based on column names:

* the latitude variable is guessed by looking for columns named lat or latitude (case-insensitive)
* the longitude variable is guessed by looking for lng, long, or longitude

You can always explicitly identify latitude/longitude columns by providing lng and lat arguments to the layer function.

For example, we do not specify the values for the arguments lat and lng in addCircles() below, but the columns Lat and Long in the data frame df will be automatically used:


```{r}
m  <-  leaflet() %>% 
        addTiles() %>% 
        setView(lat = 48.853 , lng = 2.35, zoom = 15) 
        #carte centrée sur Notre-dame de Paris
m
```


You can always explicitly identify latitude/longitude columns by providing lng and lat arguments to the layer function.

For example, we do not specify the values for the arguments lat and lng in addCircles() below, but the columns Lat and Long in the data frame df will be automatically used:




```{r}
# add some circles to a map
df <-  data.frame(Lat = 1:10, Long = rnorm(10))
leaflet(df) %>% 
        addCircles()

```

You can also explicitly specify the Lat and Long columns (see below for more info on the ~ syntax):


```{r}
leaflet(df) %>% 
        addCircles(lng = ~Long, lat = ~Lat)
```

A map layer may use a different data object to override the data provided in leaflet(). We can rewrite the above example as:

```{r}
leaflet() %>% 
        addCircles(data = df)

leaflet() %>% 
        addCircles(data = df, lat = ~ Lat, lng = ~ Long)
```


Below are examples of using sp and maps, respectively:

```{r}
library(sp)
library(rgeos)
Sr1  <-  Polygon(cbind(c(2, 4, 4, 1, 2), c(2, 3, 5, 4, 2)))
Sr2 <-  Polygon(cbind(c(5, 4, 2, 5), c(2, 3, 2, 2)))
Sr3 <-  Polygon(cbind(c(4, 4, 5, 10, 4), c(5, 3, 2, 5, 5)))
Sr4 <-  Polygon(cbind(c(5, 6, 6, 5, 5), c(4, 4, 3, 3, 4)), hole = TRUE)
Srs1 <-  Polygons(list(Sr1), "s1")
Srs2 <-  Polygons(list(Sr2), "s2")
Srs3 <-  Polygons(list(Sr4, Sr3), "s3/4")

# SpP  <-  SpatialPolygons(list(Srs1), 1:1)
# SpP  <-  SpatialPolygons(list(Srs2), 1:1)
# SpP <-  SpatialPolygons(list(Srs1, Srs2), 1:2 )
# SpP  <-  SpatialPolygons(list(Srs3), 1:1)
SpP  <-  SpatialPolygons(list(Srs1, Srs2, Srs3), 1:3)



leaflet(height = "300px") %>% 
        addPolygons(data = SpP)
```


With map
```{r}
library(maps)
mapStates = map("state", fill = TRUE, plot = FALSE)
leaflet(data = mapStates) %>% 
        addTiles() %>%
        addPolygons(fillColor = topo.colors(10, alpha = NULL), 
                    stroke = FALSE)
```


The Formula Interface
----------------------

The arguments of all layer functions can take normal R objects, such as a numeric vector for the lat argument, or a character vector of colors for the color argument. They can also take a one-sided formula, in which case the formula will be evaluated using the data argument as the environment. For example, ~ x means the variable x in the data object, and you can write arbitrary expressions on the right-hand side, e.g., ~ sqrt(x + 1).

```{r}
m = leaflet() %>% 
        addTiles()
df = data.frame(
        lat = rnorm(100),
        lng = rnorm(100),
        size = runif(100, 5, 20),
        color = sample(colors(), 100)
)
m <- leaflet(df) %>% 
        addTiles()
m %>% addCircleMarkers(radius = ~size, color = ~color, fill = FALSE)

m %>% addCircleMarkers(radius = runif(100, 4, 10), color = c('red'))
```


Using Basemaps
===============

Leaflet supports basemaps using map tiles, popularized by Google Maps and now used by nearly all interactive web maps.

Default (OpenStreetMap) Tiles
------------------------------
The easiest way to add tiles is by calling `addTiles()` with no arguments; by default, OpenStreetMap tiles are used.

Third-Party Tiles
-------------------
Alternatively, many popular free third-party basemaps can be added using the `addProviderTiles()` function, which is implemented using the [leaflet-providers plugin](https://github.com/leaflet-extras/leaflet-providers). See [here](http://leaflet-extras.github.io/leaflet-providers/preview/index.html) for the complete set.

As a convenience, leaflet also provides a named list of all the third-party tile providers that are supported by the plugin. This enables you to use auto-completion feature of your favorite R IDE (like RStudio) and not have to remember or look up supported tile providers; just type  `providers$` and choose from one of the options. You can also use `names(providers)`to view all of the options.

```{r}
m <- leaflet() %>% setView(lng = -71.0589, lat = 42.3601, zoom = 12)

# add default tiles
m %>% addTiles()

# or
m %>% addProviderTiles(providers$OpenTopoMap)

```

Note that some tile set providers require you to register; see the project page for more information. You can pass access tokens/keys, and other options, to the tile provider by populating the options argument with the providerTileOptions() function.

Custom Tile URL Template
-----------------------------
If you happen to have a custom map tile URL template to use, you can provide it as an argument to addTiles().

WMS Tiles
---------
You can use addWMSTiles() to add WMS (Web Map Service) tiles. The map below shows the Base Reflectivity (a measure of the intensity of precipitation occurring) using the WMS from the Iowa Environmental Mesonet:

```{r}
leaflet() %>% 
        addTiles() %>% 
        setView(-93.65, 42.0285, zoom = 4) %>%
        addWMSTiles(
                "http://mesonet.agron.iastate.edu/cgi-bin/wms/nexrad/n0r.cgi",
                layers = "nexrad-n0r-900913",
                options = WMSTileOptions(format = "image/png", transparent = TRUE),
                attribution = "Weather data © 2012 IEM Nexrad"
        )
```

Combining Tile Layers
----------------------
You aren’t restricted to using a single basemap on a map; you can stack them by adding multiple tile layers. This generally only makes sense if the front tiles consist of semi transparent tiles, or have an adjusted opacity via the options argument.

```{r}
m %>% addProviderTiles(providers$MtbMap) %>%
        addProviderTiles(providers$Stamen.TonerLines,
                         options = providerTileOptions(opacity = 0.35)) %>%
        addProviderTiles(providers$Stamen.TonerLabels)
```



