--- 
title: "Plotly" 
author: "Brian Caffo, notes by BFC" 
date: "4 septembre 2017" 
output: html_document
---


Installation
=============

```{r Install, options} 
# install.packages("plotly") 
library(plotly, quietly = TRUE) 
```

Basic Scatterplot
=================

A basic scatterplot is easy to make, with the added benefit of tooltips that 
appear when your mouse hovers over each point. Specify a scatterplot by 
indicating type = "scatter". Notice that the arguments for the x and y variables
as specified as formulas, with the tilde operator (~) preceding the variable
that you're plotting.


```{r}
plot_ly(mtcars, x = ~wt, y = ~mpg, type = "scatter")
```
or (no! old)

```{r, echo=TRUE}
#plot_ly(mtcars, x = wt, y = mpg, mode = "markers")
```

Scatterplot Color
=================

You can add color to your scatterplot points according to a categorical variable
in the data frame you use with `plot_ly()`.

```{r}
plot_ly(mtcars, x = ~wt, y = ~mpg, type = "scatter", color = ~factor(cyl))
```

Continuous Color
================

You can also show continuous variables with color in scatterplots.

```{r}
plot_ly(mtcars, x = ~wt, y = ~mpg, type = "scatter", color = ~disp)

```

Scatterplot Sizing
==================

By specifying the `size` argument you can make each point in your scatterplot a
different size.

```{r}
plot_ly(mtcars, x = ~wt, y = ~mpg, type = "scatter", mode = "markers", # added mode
        color = ~factor(cyl), size = ~hp)
```




3D Scatterplot
==============
You can create a three-dimensional scatterplot with the `type = "scatter3d"`
argument. If you click and drag these scatterplots you can view them from
different angles.

```{r, echo=FALSE, message=FALSE}
set.seed(2016-07-21)
temp <- rnorm(100, mean = 30, sd = 5)
pressue <- rnorm(100)
dtime <- 1:100
plot_ly(x = ~temp, y = ~pressue, z = ~dtime,
        type = "scatter3d", color = ~temp)
```


## Line Graph

Line graphs are the default graph for `plot_ly()`. They're of course useful for
showing change over time:

```{r, echo=FALSE, message=FALSE}
data("airmiles")
plot_ly(x = ~time(airmiles), y = ~airmiles, type = "scatter", mode = "lines")
```


## Multi Line Graph

You can show multiple lines by specifying the column in the data frame that
separates the lines:

```{r, echo=FALSE, message=FALSE}
library(tidyr)
library(dplyr)
data("EuStockMarkets")

stocks <- as.data.frame(EuStockMarkets) %>%
  gather(index, price) %>%
  mutate(time = rep(time(EuStockMarkets), 4))

plot_ly(stocks, x = ~time, y = ~price, color = ~index, type = "scatter", mode = "lines")
```



Histogram
==========
A histogram is great for showing how counts of data are distributed. Use the
`type = "histogram"` argument.

```{r, echo=FALSE, message=FALSE}
plot_ly(x = ~precip, type = "histogram")
```

Boxplot
=======
Boxplots are wonderful for comparing how different datasets are distributed.
Specify `type = "box"` to create a boxplot.

```{r, echo=FALSE, message=FALSE}
plot_ly(iris, y = ~Petal.Length, color = ~Species, type = "box")
```


Heatmap
========

Heatmaps are useful for displaying three dimensional data in two dimensions,
using color for the third dimension. Create a heatmap by using the
`type = "heatmap"` argument.

```{r, echo=FALSE, message=FALSE}
terrain1 <- matrix(rnorm(100*100), nrow = 100, ncol = 100)
plot_ly(z = ~terrain1, type = "heatmap")
```

## 3D Surface

Why limit yourself to two dimensions when you can render three dimensions on a
computer!? Create moveable 3D surfaces with `type = "surface"`.

```{r, echo=FALSE, message=FALSE}
terrain2 <- matrix(sort(rnorm(100*100)), nrow = 100, ncol = 100)
plot_ly(z = ~terrain2, type = "surface")
```


Choropleth Maps
===============

```{r, echo=FALSE, message=FALSE}
# Setup
# -----
# Create data frame
state_pop <- data.frame(State = state.abb, Pop = as.vector(state.x77[,1]))
# Create hover text
state_pop$hover <- with(state_pop, paste(State, '<br>', "Population:", Pop))
# Make state borders white
borders <- list(color = toRGB("red"))
# Set up some mapping options
map_options <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

# Map
# ---

plot_ly(z = ~state_pop$Pop, text = ~state_pop$hover, locations = ~state_pop$State, 
        type = 'choropleth', locationmode = 'USA-states', 
        color = state_pop$Pop, colors = 'Blues', marker = list(line = borders)) %>%
  layout(title = 'US Population in 1975', geo = map_options)
```

ggplot

```{r}
library(ggplot2, quietly = TRUE)

# simple example
ggiris <- qplot(Petal.Width, Sepal.Length, data = iris, color = Species)
ggly <- ggplotly(ggiris)
ggly


data(canada.cities, package = "maps")
viz <- ggplot(canada.cities, aes(long, lat)) +
  borders(regions = "canada") +
  coord_equal() +
  geom_point(aes(text = name, size = pop), colour = "red", alpha = 1/2)
ggplotly(viz, tooltip = c("text", "size"))


# # highlighting lines
# demo("highlight-ggplotly", package = "plotly")

# # client-side linked brushing
# library(crosstalk)
# d <- SharedData$new(mtcars)
# subplot(
#  qplot(data = d, x = mpg, y = wt),
#  qplot(data = d, x = mpg, y = vs)
# )
# 
# # client-side linked brushing in a scatterplot matrix
# SharedData$new(iris) %>%
#   GGally::ggpairs(aes(colour = Species), columns = 1:4) %>%
#   ggplotly(tooltip = c("x", "y", "colour"))

```



Resources
=========

- [The Plolty Website](https://plot.ly/)
- [The Plotly R API](https://plot.ly/r/)
- [The Plotly R Package on GitHub](https://github.com/ropensci/plotly)
- [The Plotly R Cheatsheet](https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf)
- ["Plotly for R" book by Carson Sievert](https://cpsievert.github.io/plotly_book/)

